Picky with peakpicking: assessing chromatographic peak quality with simple

### 2023-08-02 manuscript/manuscript.Rmd
Renaming supp fig legend txt file
Combining supplemental tables into single Excel file
Also contains titles and legends

```{r full model estimate csv write}
set.seed(123)
both_datasets <- list(MESOSCOPE=MS3000_features, Falkor=FT2040_features) %>%
  bind_rows(.id = "dataset") %>%
  filter(feat_class%in%c("Good", "Bad")) %>%
  mutate(feat_bin=ifelse(feat_class=="Good", 1, 0))

glm_estimates <- both_datasets %>%
  select(feat_bin, all_of(all_params)) %>%
  # mutate(across(-feat_bin, \(x)scale(x)[,1])) %>%
  glm(formula = feat_bin~., family=binomial) %>%
  broom::tidy() %>%
  rename_with(~paste0("GLM.", .x), -term)

glmnet_estimates <- both_datasets %>% 
  select(all_of(all_params), c("max_SNR", "max_cor", "medtop3_cor", "log_med_cor")) %>% 
  data.matrix() %>%
  glmnet::cv.glmnet(y=both_datasets$feat_bin, family="binomial", alpha=0.5) %>%
  coef() %>%
  as.matrix() %>%
  as.data.frame() %>%
  rownames_to_column("term") %>%
  dplyr::rename(`Elastic.net.estimate`=s1)

rf_estimates <- both_datasets %>% 
  select(all_of(all_params), c("max_SNR", "max_cor", "medtop3_cor", "log_med_cor")) %>% 
  randomForest::randomForest(factor(both_datasets$feat_bin), importance=TRUE) %>%
  pluck("importance") %>%
  as.data.frame() %>%
  rownames_to_column("term") %>%
  select(term, RF.accuracy.decrease=MeanDecreaseAccuracy,
         RF.Gini.Decrease=MeanDecreaseGini)

param_longnames %>%
  as.data.frame() %>%
  set_names("long_name") %>%
  rownames_to_column("term") %>%
  left_join(glm_estimates, by="term") %>%
  left_join(glmnet_estimates, by="term") %>%
  left_join(rf_estimates, by="term") %>%
  write_csv("manuscript/supplement/supp_table_1_full_model_estimates.csv")
```

@@

writeLines(comb_supp_text, "manuscript/supplement/supp_text_to_copy_paste.txt")
```

@@

```{r confusion matrices for all the models for supplement}
model_levels <- c("all_params", "xcms_params", "raw_data_params", 
                  "ridge", "elastic", "lasso", "rf")
model_labels <- c("Full model", "XCMS model", "Two-parameter model",
                  "Ridge regression", "Elastic net", "Lasso regression", 
                  "Random forest")
fdr_gff_individual <- fdr_gff_cross_train_df %>%
  bind_rows(fdr_gff_cross_train_df_fancy %>% dplyr::rename(which_params=which_alpha)) %>%
  separate("train_test", into = c("train", "test"), sep = "-") %>%
  select(-`NA`) %>%
  select(model_type=which_params, train, test, TP, TN, FP, FN, FDR, GFF) %>%
  mutate(model_type=factor(model_type, levels=model_levels, labels=model_labels)) %>%
  arrange(model_type)
fdr_gff_full_model <- bind_rows(lst(FT2040_features, MS3000_features), .id = "dataset") %>%
  mutate(pred_prob=predict(object=both_min_model, newdata = ., type = "response")) %>%
  mutate(pred_class=ifelse(pred_prob>0.5, "Good", "Bad")) %>%
  mutate(error_type=case_when(
    feat_class=="Bad" & pred_class=="Bad" ~ "TN",
    feat_class=="Bad" & pred_class=="Good" ~ "FP",
    feat_class=="Good" & pred_class=="Bad" ~ "FN",
    feat_class=="Good" & pred_class=="Good" ~ "TP"
  )) %>%
  drop_na() %>%
  bind_rows(combined=., each=., .id = "calc_type") %>% 
  mutate(calc_type=ifelse(calc_type=="each", dataset, calc_type)) %>%
  select(-dataset) %>%
  group_by(calc_type) %>%
  count(error_type) %>%
  ungroup() %>%
  pivot_wider(names_from=error_type, values_from = n, values_fill = 0) %>%
  mutate(FDR=FP/(FP+TP)) %>%
  mutate(GFF=TP/(FN+TP)) %>%
  mutate(model_type="Two-parameter model") %>%
  mutate(train="Both") %>%
  mutate(test=case_when(
    calc_type=="FT2040_features"~"FT2040",
    calc_type=="MS3000_features"~"MS3000",
    calc_type=="combined"~"Both"
  )) %>%
  select(model_type, train, test, TP, TN, FP, FN, FDR, GFF)

fdr_gff_strict_all <- bind_rows(lst(FT2040_features, MS3000_features), .id = "dataset") %>%
  mutate(pred_prob=predict(object=both_min_model, newdata = ., type = "response")) %>%
  mutate(pred_class=ifelse(pred_prob>0.5, "Good", "Bad")) %>%
  mutate(error_type=case_when(
    feat_class=="Bad" & pred_class=="Bad" ~ "TN",
    feat_class=="Bad" & pred_class=="Good" ~ "FP",
    feat_class=="Good" & pred_class=="Bad" ~ "FN",
    feat_class=="Good" & pred_class=="Good" ~ "TP"
  )) %>%
  drop_na() %>%
  bind_rows(combined=., each=., .id = "calc_type") %>% 
  mutate(calc_type=ifelse(calc_type=="each", dataset, calc_type)) %>%
  select(-dataset) %>%
  group_by(calc_type) %>%
  count(error_type) %>%
  ungroup() %>%
  pivot_wider(names_from=error_type, values_from = n, values_fill = 0) %>%
  mutate(FDR=FP/(FP+TP)) %>%
  mutate(GFF=TP/(FN+TP)) %>%
  mutate(model_type="Two-parameter model") %>%
  mutate(train="Both") %>%
  mutate(test=case_when(
    calc_type=="FT2040_features"~"FT2040",
    calc_type=="MS3000_features"~"MS3000",
    calc_type=="combined"~"Both"
  )) %>%
  select(model_type, train, test, TP, TN, FP, FN, FDR, GFF)

fdr_gff_stricter_thresh <- error_type_all_datasets %>%
  set_names(c("test", "FN", "FP", "TN", "TP", "FDR", "GFF")) %>%
  mutate(model_type="Both (threshold 0.9)") %>%
  mutate(train="Both") %>%
  select(model_type, train, test, TP, TN, FP, FN, FDR, GFF)

bind_rows(fdr_gff_individual, fdr_gff_full_model, fdr_gff_stricter_thresh) %>%
  write_csv("manuscript/supplement/supp_table_2_model_confusion_matrices.csv")
```

>>>>

@@

writeLines(comb_supp_text, "manuscript/supplement/supp_fig_legends.txt")
```

@@

```{r supp table Excel write model params and confusion matrices}
library(openxlsx)
wb <- createWorkbook(creator = "William Kumler, wkumler@uw.edu")
addWorksheet(wb, sheetName = "Table titles and legends")
data.frame("Supplemental table titles and legends"=c(
  "Supplemental table 1:",
  "Model parameters and significance according to both logistic and random forest models",
  "Model outputs from the full (all parameters) model trained on the combined Falkor",
  "and MESOSCOPE datasets. The parameter name is reported alongside its more detailed",
  "name, with the logistic model estimate, standard error, test statistic, and p-value",
  "for each predictive term. We also included the estimate obtained from the elastic",
  "net model (alpha=0.5) and the accuracy and Gini index decreases per term. NA values",
  "are reported for several terms that were removed from the non-regularized regression",
  "because they were too highly correlated with other parameters.",
  "",
  "Supplemental table 2:",
  "Confusion matrices for logistic, regularized, and random forest regressions",
  "Confusion matrices reported in long format for the different models reported in this",
  "manuscript. Models were all both internally tested using the same dataset for testing",
  "and training as well as cross-validated against the other fully labeled dataset. For",
  "the two-parameter model, we also report the values obtained when training on the two",
  "datasets combined (train = Both). For each model and test-train set we report the",
  'absolute number of false positives (manually labeled as "Bad" but model labeled as',
  '"Good"), false negatives (manually labeled as "Good" but model labeled as "Bad"),',
  'true positives (both manual and model labelled as "Good"), and true negatives (both',
  'manual and model labeled as "Bad") as well as the calculated false discovery rate',
  "(FDR, calculated as FP/(FP+TP)) and the fraction of good features found (GFF,",
  "calculated as TP/(TP+FN)). Unless otherwise specified, all values reported here",
  "used a likelihood threshold of 0.5 for categorization."
), check.names = FALSE) %>%
  writeData(wb = wb, sheet = "Table titles and legends")

addWorksheet(wb, sheetName = "Supp table 1 Model parameters")
set.seed(123)
both_datasets <- list(MESOSCOPE=MS3000_features, Falkor=FT2040_features) %>%
  bind_rows(.id = "dataset") %>%
  filter(feat_class%in%c("Good", "Bad")) %>%
  mutate(feat_bin=ifelse(feat_class=="Good", 1, 0))
glm_estimates <- both_datasets %>%
  select(feat_bin, all_of(all_params)) %>%
  # mutate(across(-feat_bin, \(x)scale(x)[,1])) %>%
  glm(formula = feat_bin~., family=binomial) %>%
  broom::tidy() %>%
  rename_with(~paste0("GLM.", .x), -term)
glmnet_estimates <- both_datasets %>% 
  select(all_of(all_params), c("max_SNR", "max_cor", "medtop3_cor", "log_med_cor")) %>% 
  data.matrix() %>%
  glmnet::cv.glmnet(y=both_datasets$feat_bin, family="binomial", alpha=0.5) %>%
  coef() %>%
  as.matrix() %>%
  as.data.frame() %>%
  rownames_to_column("term") %>%
  dplyr::rename(`Elastic.net.estimate`=s1)
rf_estimates <- both_datasets %>% 
  select(all_of(all_params), c("max_SNR", "max_cor", "medtop3_cor", "log_med_cor")) %>% 
  randomForest::randomForest(factor(both_datasets$feat_bin), importance=TRUE) %>%
  pluck("importance") %>%
  as.data.frame() %>%
  rownames_to_column("term") %>%
  select(term, RF.accuracy.decrease=MeanDecreaseAccuracy,
         RF.Gini.Decrease=MeanDecreaseGini)
param_longnames %>%
  as.data.frame() %>%
  set_names("long_name") %>%
  rownames_to_column("term") %>%
  left_join(glm_estimates, by="term") %>%
  left_join(glmnet_estimates, by="term") %>%
  left_join(rf_estimates, by="term") %>%
  writeData(wb = wb, sheet = "Supp table 1 Model parameters")

addWorksheet(wb, sheetName = "Supp table 2 Confusion matrices")
model_levels <- c("all_params", "xcms_params", "raw_data_params", 
                  "ridge", "elastic", "lasso", "rf")
model_labels <- c("Full model", "XCMS model", "Two-parameter model",
                  "Ridge regression", "Elastic net", "Lasso regression", 
                  "Random forest")
fdr_gff_individual <- fdr_gff_cross_train_df %>%
  bind_rows(fdr_gff_cross_train_df_fancy %>% dplyr::rename(which_params=which_alpha)) %>%
  separate("train_test", into = c("train", "test"), sep = "-") %>%
  select(-`NA`) %>%
  select(model_type=which_params, train, test, TP, TN, FP, FN, FDR, GFF) %>%
  mutate(model_type=factor(model_type, levels=model_levels, labels=model_labels)) %>%
  arrange(model_type)
fdr_gff_full_model <- bind_rows(lst(FT2040_features, MS3000_features), .id = "dataset") %>%
  mutate(pred_prob=predict(object=both_min_model, newdata = ., type = "response")) %>%
  mutate(pred_class=ifelse(pred_prob>0.5, "Good", "Bad")) %>%
  mutate(error_type=case_when(
    feat_class=="Bad" & pred_class=="Bad" ~ "TN",
    feat_class=="Bad" & pred_class=="Good" ~ "FP",
    feat_class=="Good" & pred_class=="Bad" ~ "FN",
    feat_class=="Good" & pred_class=="Good" ~ "TP"
  )) %>%
  drop_na() %>%
  bind_rows(combined=., each=., .id = "calc_type") %>% 
  mutate(calc_type=ifelse(calc_type=="each", dataset, calc_type)) %>%
  select(-dataset) %>%
  group_by(calc_type) %>%
  count(error_type) %>%
  ungroup() %>%
  pivot_wider(names_from=error_type, values_from = n, values_fill = 0) %>%
  mutate(FDR=FP/(FP+TP)) %>%
  mutate(GFF=TP/(FN+TP)) %>%
  mutate(model_type="Two-parameter model") %>%
  mutate(train="Both") %>%
  mutate(test=case_when(
    calc_type=="FT2040_features"~"FT2040",
    calc_type=="MS3000_features"~"MS3000",
    calc_type=="combined"~"Both"
  )) %>%
  select(model_type, train, test, TP, TN, FP, FN, FDR, GFF)
fdr_gff_strict_all <- bind_rows(lst(FT2040_features, MS3000_features), .id = "dataset") %>%
  mutate(pred_prob=predict(object=both_min_model, newdata = ., type = "response")) %>%
  mutate(pred_class=ifelse(pred_prob>0.5, "Good", "Bad")) %>%
  mutate(error_type=case_when(
    feat_class=="Bad" & pred_class=="Bad" ~ "TN",
    feat_class=="Bad" & pred_class=="Good" ~ "FP",
    feat_class=="Good" & pred_class=="Bad" ~ "FN",
    feat_class=="Good" & pred_class=="Good" ~ "TP"
  )) %>%
  drop_na() %>%
  bind_rows(combined=., each=., .id = "calc_type") %>% 
  mutate(calc_type=ifelse(calc_type=="each", dataset, calc_type)) %>%
  select(-dataset) %>%
  group_by(calc_type) %>%
  count(error_type) %>%
  ungroup() %>%
  pivot_wider(names_from=error_type, values_from = n, values_fill = 0) %>%
  mutate(FDR=FP/(FP+TP)) %>%
  mutate(GFF=TP/(FN+TP)) %>%
  mutate(model_type="Two-parameter model") %>%
  mutate(train="Both") %>%
  mutate(test=case_when(
    calc_type=="FT2040_features"~"FT2040",
    calc_type=="MS3000_features"~"MS3000",
    calc_type=="combined"~"Both"
  )) %>%
  select(model_type, train, test, TP, TN, FP, FN, FDR, GFF)
fdr_gff_stricter_thresh <- error_type_all_datasets %>%
  set_names(c("test", "FN", "FP", "TN", "TP", "FDR", "GFF")) %>%
  mutate(model_type="Both (threshold 0.9)") %>%
  mutate(train="Both") %>%
  select(model_type, train, test, TP, TN, FP, FN, FDR, GFF)
bind_rows(fdr_gff_individual, fdr_gff_full_model, fdr_gff_stricter_thresh) %>%
  writeData(wb = wb, sheet = "Supp table 2 Confusion matrices")
saveWorkbook(wb, file = "manuscript/supplement/BMC_supp_tables.xlsx", overwrite = TRUE)
```
